Data Specialization for a Staged Lambda Calculus

### 2014-11-08 paper/implementation.tex
small edit to intro of implementation

The rules of \cref{sec:splitting} can be optimized in many cases.
For instance, multiple trivial subterm precomputations can be merged:
\[
\infer {\splittwo {(e_1,e_2)} A {p,l.(r_1,r_2)}}{\splittwo {e_1} A {(),\_.r_1} & \splittwo {e_2} A {p,l.r_2}}
\]
This sort of optimization is easy, because it is local on the structure of $e$,
but the problem becomes more difficult when trying to optimize across functions, including recursive calls.
For example, consider the stage \bbtwo\ term,
\begin{lstlisting}
2`prev{
  1`letfun fact (n : int) : int = 
    if n <= 0 then 1 else fact(n-1)*n
  in next{`2`hold {`1`fact 5`2`}-100`1`}`2`
}`
\end{lstlisting}
Noting that the precomputation of the body of \texttt{fact} could be given the type,
\begin{lstlisting}
datatype prec = L | R of prec
\end{lstlisting}
the whole expression splits into:
\begin{lstlisting}
1`letfun fact (n : int) : int * prec = 
  if n <= 0 then (1,L) 
  else let (y,z) = fact(n-1) in (y*n,R z)
in fact 5`

>>>>

Applying the rules of \cref{sec:splitting} naively can yield inefficient split programs.
In many cases, the results can be improved by considering special cases in individual rules.
For instance, trivial subterm precomputations needn't be saved at tuples:
\[
\infer {\splittwo {(e_1,e_2)} A {p,l.(r_1,r_2)}}{\splittwo {e_1} A {(),\_.r_1} & \splittwo {e_2} A {p,l.r_2}}
\]
The problem becomes more difficult when trying to optimize across functions, including recursive calls.
For example, consider the stage \bbtwo\ term,
\begin{lstlisting}
2`prev{
  1`letfun fact (n : int) : int = 
    if n <= 0 then 1 else fact(n-1)*n
  in next{`2`hold {`1`fact 5`2`}-100`1`}`2`
}`
\end{lstlisting}
Noting that the precomputation of the body of \texttt{fact} could be given the type,
\begin{lstlisting}
datatype prec = L | R of prec
\end{lstlisting}
the whole expression splits into:
\begin{lstlisting}
1`letfun fact (n : int) : int * prec = 
  if n <= 0 then (1,L) 
  else let (y,z) = fact(n-1) in (y*n,R z)
in fact 5`
